---
title: "Project 4"
author: "Yue Peng, Kun Meng, Ludan Zhang, Jiachen Zhang"
date: "Due October 7, 2016 at 11:59pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:

The goal of this project is to get aquainted with the Rstudio server as well as to work on creating functions. You will do some function refactoring as well.[Lab 4 - Markdown]()


# Part I: Function Building with the jackknife function


## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance

<small>(Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)</small>



Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.


```{r, eval=FALSE}
mean.jackknife <- function(a_vector) {
  # Put your function in here
  n <- length(a_vector)
  cita <- rep(NA, n) 
  for (i in 1:n){
    cita[i] <- mean(a_vector[-i])
  }
  jackknife.stderr <- sqrt((n-1)^2/n * var(cita))
  return(jackknife.stderr)
}
```



#### Jackknife for the mean Example: test your code here


```{r, eval = FALSE}
some_normals <- rnorm(100,mean=7,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals)/sqrt(length(some_normals)))
all.equal(formula_se_of_mean,mean.jackknife(some_normals))
```



##Jackknife for Gamma Parameters


Recall our friend the method of moments estimator:

```{r, eval=FALSE}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2/v
  s <- v/m
  return(c(a=a,s=s))
}
```


###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r, eval=FALSE }
gamma.jackknife <- function(a_vector) {
  n <- length(a_vector)
  gamma_est <- matrix(rep(NA, n), ncol = n,nrow = 2)
  for (i in 1:n){
    gamma_est[, i] <- gamma.est(a_vector[-i])
  }
  jackknife.stderrs <- c(NA, NA)
  jackknife.stderrs[1] <- sqrt((n-1)^2/n*var(gamma_est[1,]))
  jackknife.stderrs[2] <- sqrt((n-1)^2/n*var(gamma_est[2,]))
  names(jackknife.stderrs) <- c("a", "s" )
  return(jackknife.stderrs)
}
```



###Jackknife for Gamma Parameters Example


```{r, eval=FALSE}
input <- rgamma(1000, shape=0.1, scale=10)
gamma.est(input)
gamma.jackknife(input)
```



##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r, eval=FALSE}
jackknife.lm <- function(df, formula, p) {
  n <- nrow(df)
  jackknife.ests <- matrix(NA, ncol = n, nrow = p)
  name1 <- unlist(strsplit(formula, "~"))[1]
  name2 <- unlist(strsplit(formula, "~"))[2]
  formula1 <- paste(name1,"~", name2, sep = "")
  for (i in 1:n){
    my_lm <- lm(formula1, data = df[-i,])
    jackknife.ests[,i] <- coefficients(my_lm)
  }
  jackknife.stderr <- vector(length = p)
  for (i in 1:p){
  jackknife.stderr[i] <- sqrt((n-1)^2/n*var(jackknife.ests[i,]))
  }
  return(jackknife.stderr)
}
```



###Jackknife for linear regression coefficients Example


```{r, eval=FALSE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
jackknife.lm(df=data,formula="output~input",p=2)
```




##Refactoring the Jackknife


- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```

- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation



###The Common Operation


- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:



## Goal: 


- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 
    ```{r, eval=FALSE}
          omit.case <- function(the_data,omitted_point) {
        # This should take the data and omit one point at a time and return the new data
          ifelse(is.vector(the_data), x <- the_data[-omitted_point], x <- the_data[-omitted_point,]) 
          return(x)
            }
    ```
- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does.  
    ```{r,eval=FALSE}



```{r, eval=FALSE}

jackknife <- function(estimator,the_data) {
      
        omit_and_est <- function(omit) {
              # This function should take the output of omit.case and use it as input for the estimator
          estimator(omit.case(the_data, omit))
        }
  
  # Define n to be lenth or number of rows here
  n <- ifelse(is.vector(the_data),  length(the_data), nrow(the_data)) 
  
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



---

###It works


```{r, eval=FALSE}
jackknife(estimator = mean, the_data = some_normals)
all.equal(jackknife(estimator=mean,the_data=some_normals),
          mean.jackknife(some_normals))
```

---



```{r, eval=FALSE}
all.equal(jackknife(estimator = gamma.est, the_data = data$input),
          gamma.jackknife(data$input))
```

---


```{r, eval=FALSE}
all.equal(jackknife(estimator = gamma.est,the_data = data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```


```{r, eval=FALSE}
est.coefs <- function(the_data) {
  return(lm(output~input,data=the_data)$coefficients)
}
est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---


```{r, eval=FALSE}
jackknife(estimator=est.coefs,the_data=data)
all.equal(jackknife(estimator=est.coefs,the_data=data),
          jackknife.lm(df=data,formula="output~input",p=2))
```


##Further Refactoring of jackknife()


The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:

    ```{r, eval=FALSE}
      data_size <- function(the_data) {
          n <- ifelse(is.vector(the_data),  length(the_data), nrow(the_data)) 
      }
    ```
- Write a funcrion that calculate the variance of all the estimates and returns the standard error
```{r, eval=FALSE}
scale_and_sqrt_vars <- function(jackknife.ests,n) {
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



###Now invoke those functions

```{r, eval=FALSE}
jackknife <- function(estimator,the_data) {
  n <- data_size(the_data)
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data,omit))
  }
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  return(scale_and_sqrt_vars(jackknife.ests,n))
}
```

