---
title: "Project 7"
author: "Yue Peng, Ludan Zhang, Jiachen Zhang, Kun Meng"
date: "Due October 26, 2016 at 5:00pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:

The goal of this project is to get better  aquainted with Dplyr and working with data. 

[Project 7 Markdown]()

# New York City Flights 13

- `flights` [336,776 x 16]. Every flight leaving NYC in 2013.
- `weather` [ 8719 x 14]. Hourly Weather Data.
- `planes` [3322 x 9]. Plane data.
- `airports` [1397 x 7]. Airport data. 


```{R}
library(nycflights13)
library(dplyr)
flights
```


# Basic Dyplyr

The deplyr package uses verbs to perform data manipulation.
- `filter()` and `slice()`
- `arrange()`
- `select()` and `rename()`
- `distinct()`
- `mutate()` and `transmute()`
- `sample_n()` and `sample_frac()`

## Filtering Rows

1. Filter all flights on your birthday. 
```{r}
bir_month <- 8
bir_day <- 14
flights %>%
filter(month == bir_month, day == bir_day)
```

2. Filter all flights in rows 567 - 589
```{r}
flights %>%
  slice(567:589)
```

## Arranging Rows

1. Use  `arrange()` to order the data by `year, month and day`
```{r}
flights %>%
  arrange(year, month, day)
```
2. Use `desc()` to arrange flights by arrival delay. 
```{r}
flights %>%
  arrange(desc(arr_delay))
```

## Select

1. Select columns containing date information.
```{r}
flights %>%
  select(year, month, day, time_hour)
```

2. Select columns including only plane information.
```{r}
flights %>%
  select(carrier, flight, tailnum)
```

3. Rename `tailnum` to `tail_num`
```{r}
flights %>%
  select(tail_num = tailnum)
#rename(flights, tail_num = tailnum)
```

## Mutating Rows

1. Create a variable called `gain` which describes the difference between arrival delap and departure delay.
```{r}
flights %>%
  mutate(gain = arr_delay - dep_delay)
```

2. Create a variable called speed which is distance over time in hours. 
```{r}
flights %>%
  mutate(speed = distance / air_time * 60)
```

## Summarise

1. Summarise the average delay for all flights.
```{r}
flights %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE))
```

# Overall Questions 

1. Split the complete dataset into individual planes and then summarise each plane by counting the number of flights (count = n()) and computing the average distance and arrival delay .  Then plot this:
```{r}
delay <- flights %>%
  group_by(tailnum) %>%
  summarise(count = n(), dist = mean(distance, na.rm = TRUE), delay = mean(arr_delay, na.rm = TRUE))

```

```{r}
library(ggplot2)
ggplot(delay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area()
```


2.  Find the number of planes and the number of flights that go to each possible destination. 
```{r}
flights %>%
  group_by(dest) %>%
  summarise(planes = n_distinct(tailnum), flights = n())
```


### More Difficult

In the videos we learned about piping our commands. The goal of piping is to help use with the flow of our analysis. The steps we take are:
- Think about it. 
- Describe it. 
- Do it. 


### Consider the function below:



```{r}
bas_r <- filter(
      summarise(
        select(
          group_by(flights, year, month, day),
          arr_delay, dep_delay
        ),
        arr = mean(arr_delay, na.rm = TRUE),
        dep = mean(dep_delay, na.rm = TRUE)
      ),
      arr > 30 | dep > 30
    )

```



  a. What are some problems with this function? 
```{r}
# If you want to understand this chunk of codes, you must read from inside to outside. It is not readable.
```
  b. How easy is it to follow the logic of this? 
```{r}
# Not easy to read in this way. We can rewrite it step by step, write down every step by natural thinking order.
```
  c. Rewrite this using piping to make it more understandable. 
```{r}
pip_r <- flights %>%
  group_by(year, month, day) %>%
  select(arr_delay, dep_delay) %>%
  summarise(arr = mean(arr_delay, na.rm = TRUE), dep = mean(dep_delay, na.rm = TRUE)) %>%
  filter(arr >30 | dep >30)
```
  d. Use your function testing skills to verify they return the same results. 
```{r}
all_equal(bas_r, pip_r)
```
  e.  How would you write the above in terms of SQL?
```{r}
library(RMySQL)
library(DBI)
con <-dbConnect(MySQL(),
                user = "py0814",
                password = "seven=2+5",
                host = "yuedb.ccxrmxtuzhfi.us-east-1.rds.amazonaws.com",
                dbname = "Mydata")

query <- "SELECT year, month, day, AVG(arr_delay) AS arr, AVG(dep_delay) AS dep
          FROM flights 
          GROUP BY year, month, day
          HAVING arr > 30 OR dep >30;"

dbGetQuery(con, query)
```
  
### Consider the function below



```{r}
    hourly_delay <- filter(
              summarise(
                  group_by(
                      filter(
                        flights, 
                        !is.na(dep_delay)
                        ),
                        month, day, hour
                        ),
                        delay=mean(dep_delay),
                        n=n()
                        ),
                        n>10
                        )
                                          
```


  a. What are some problems with this function? 
```{r}
# If you want to understand this chunk of codes, you must read from inside to outside. It is not readable from our natural logic.
```
  b. How easy is it to follow the logic of this? 
```{r}
# Not easy to read in this way. We can rewrite it step by step, write down every step by natural thinking order.
```
  c. Rewrite this using piping to make it more understandable. 
```{r}
hourly_delay2 <- flights %>%
  filter(!is.na(dep_delay)) %>%
  group_by(month, day, hour) %>%
  summarise(delay = mean(dep_delay), n = n()) %>%
  filter(n > 10)
```

  d. Use your function testing skills to verify they return the same results. 
```{r}
all_equal(hourly_delay, hourly_delay2)
```
  e.  How would you write the above in terms of SQL?
```{r}


query2 <- "SELECT month, day, hour, AVG(dep_delay) AS delay, COUNT(*) AS n
           FROM flights 
           WHERE dep_delay IS NOT NULL
           GROUP BY month, day, hour
           HAVING n > 10;"

dbGetQuery(con, query2)
```


## Consider how to join information from the other datasets in the package with the current information. Is there a way to tell if weather affected the departure  time? Ask yourself other questions about this data and then answer them below. 
```
We can join the flights dataset with a weather dataset which records the weather everyday in year 2013. That is to say, we can have a weather dataset containing date(primary key) and weather columns and use inner_join to join the flights dataset with the weather dataset to a new dataset called flights_weather. Then we can summarise the mean departure time of the flights for each type of weather by using code like: summarise(group_by(flights_weather,weather),mean(dep_time)). The summary would tell us if the weather affected the departure time.
```