---
title: "Project 6: part b"
author: "YUE PENG"
date: "April 14, 2016"
output:
  html_document
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(eval=FALSE)

```

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>

####[Lab 7 Markdown]()

## Setup 

For the setup of this, I have created a database called titanic. The goal will be to use this lab to work on the titanic dataset. The code below will let you connect to the database. 

```{R}
library(DBI)
library(RMySQL)
con <- dbConnect(MySQL(),
                 user = 'trial',
                 password = 'password',
                 host = '127.0.0.1',
                 dbname='titanic')

dbGetInfo(con)
dbListTables(con)
```


1. What tables are in this database?
```{r}
query <- "show tables;"
dbGetQuery(con, query)
```

2. Load the titanic data:

```{r}
titanic <- read.table("/home/php2560/titanic.csv", header=TRUE, sep=",")
```

3. Create a table in MySQL from this data. Name it `titanic_yourlastname`:

```{r}
dbWriteTable(con, "titanic_peng", titanic)
```

4. Now what tables are in this database?
```{r}
query <- "show tables;"
dbGetQuery(con, query)
```

## Basic usage of commands

We will start to use a list of commands that would be commonly used in MySQL:

These are:

- show tables;
- describe titanic_yourlastname;
- select name, pclass, age, sex from titanic limit 10;
 


We can run any of these commands as follow:

```
query <- "show tables;"
dbGetQuery(con, query)
```

Try these commands and test this out. 


## R and MySQL

5. Run the following SQL command. Before doing so describe what this is asking
```{r}
query5 <- "select pclass, survived, avg(age) as avg_age from titanic_peng
   group by pclass, survived;"
dbGetQuery(con,query5)
```


6.  Run the following SQL command. Before doing so describe what this is asking
```{r}
query6 <- "select pclass, survived, avg(age) as avg_age from titanic_peng
   where survived = 1
   group by pclass, survived;"
dbGetQuery(con, query6)
```


7. Run the following SQL command. Before doing so describe what this is asking
```{r}
query7 <- "select name, pclass, survived, age from titanic_peng 
   where name regexp '^.sen.+';"
dbGetQuery(con, query7)
```

8. Run the following SQL command. Before doing so describe what this is asking
```{r}
query8 <- "select name, pclass, survived, age from titanic_yourlastname 
   where name regexp 'Jakob.*$';"
dbGetQuery(con, query8)
```

9.  Run the following SQL commands. What is different about them? What do they return?
```{r}
query9a <- "select name, pclass, survived, age from titanic_peng 
   where name regexp 'Ivar.*$';"
dbGetQuery(con, query9a)
```
   
```{r}
query9b <- "select name, pclass, survived, age from titanic_peng 
   where name regexp ',.*Ivar.*$';"
dbGetQuery(con, query9b)
```


10. We can also plot data from this:
```{r}
myQuery <- "select pclass, avg(fare) as avg_fare from titanic_peng
              group by pclass;"
myData <- dbGetQuery(con, myQuery)
library(ggplot2)
ggplot(myData, aes(pclass, avg_fare)) + geom_bar(stat="identity")
```



## More challenging

11. Create a data frame with:
- `pclass`
- `sex`
- `age`
- `fare`
- `survived`
- `parch`
Only for people who did not survive. 
```{r}
query11 <- "select pclass, sex, age, fare, survived, parch from titanic_peng 
   where survived != 1;"
notsurvive <- dbGetQuery(con, query11)
notsurvive
```

12. Create a graph of the average survival over the different classes. Make two distinct lines for male and female passengers. Create this plot using ggplot. 
```{r}
query12 <- "select pclass, sex, avg(survived) as ave_sur from titanic_peng 
   group by pclass, sex;"
data12 <- dbGetQuery(con, query12)
data12 <- data12[2:7,]
library(ggplot2)
ggplot(data12,aes(pclass,ave_sur, color=sex, group=sex))+geom_point(size=3)+geom_line()

```
```{r}
query <- "describe titanic;"
dbGetQuery(con, query)
```
