---
title: "Project 5: Group B"
author: "October 12, 2016"
date: "Due October 12, 2016 at 5:00pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:


With this project we will be be working in small groups and on 2 different teams. 

From 1:00 to 1:20pm meet as large groups (***A*** and ***B***), read through the problem and make sure everyone understands it. Set up different tasks to complete as small groups. It may help to have multiple groups work on the same task to see if they have similar codes. 

From 1:20 to 2:30 pm work in your small group to complete your tasks. During this time you also will want to have large group meetings to bring together code. Test the code out and assign new tasks. It may help 


From 2:30 - 3:15 wrap up the simulations for all of your problems. Make sure you have running code and that it works correctly. Clean your code up and get ready to present it. 

Rest of class: Group A and Group B will both present their solutions to the class. 

[Lab 5 Group B - Markdown]()


 1. **Gambler's Ruin**: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:
    a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet. 
```{r}
busted <- function(n){
      bankroll <- 1000
      bet <- 100
      ntimes <- n
  while (bankroll>0){
    ntimes <- ntimes + 1
    ifelse(rbinom(1,1,0.5), bankroll <- bankroll + bet, bankroll <- bankroll - bet)
    if (ntimes > 5000) break
  }
   return(ntimes)
}

sim_busted <- function(placetime, n){
  
  x <- matrix(rep(0,n), ncol = n)
  result <- apply(x, 2, busted)
  return(length(result[result < placetime])/n)
}
sim_busted(100, 10000)
```
    
    b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly. 
```{r}
sim_busted(500, 10000)
```

    c. the mean time you go bust, given that you go bust within the first 5000 hands.
```{r}

sim_fivethousand_busted <- function(n){
  x <- matrix(rep(0, n), ncol = n)
  result <- apply(x, 2, busted)
  return(mean(result[result < 5000]))
}
sim_fivethousand_busted(10000)
```

    d. the mean and variance of your bankroll after 100 hands (including busts).
```{r}
bank_roll <- function(x){
  bankroll <- 1000
  bet <- 100
  for (i in 1:x){
    ifelse(rbinom(1, 1, 0.5), bankroll <- bankroll - bet, bankroll <- bankroll + bet)
    if (bankroll == 0) break
  }
  return(bankroll)
}

sim_bankroll <- function(n, hands){
x <- matrix(rep(hands, n), ncol = n)
result <- apply(x, 2, bank_roll)
return(c(mean(result), var(result)))
}
sim_bankroll(5000, 100)
```

    e. the mean and variance of your bankroll after 500 hands (including busts).
```{r}
sim_bankroll(5000, 500)
```
 
Note: you *must* stop playing if your player has gone bust. How will you handle this in the `for` loop?




2. Repeat the previous problem with betting on black in American roulette, where the probability of winning on any spin is 18/38 for an even payout.
```{r}
busted2 <- function(n){
      bankroll <- 1000
      bet <- 100
      ntimes <- n
  while (bankroll>0){
    ntimes <- ntimes + 1
    ifelse(rbinom(1, 1, 18/38), bankroll <- bankroll + bet, bankroll <- bankroll - bet)
    if (ntimes > 5000) break
  }
   return(ntimes)
}

sim_busted2 <- function(placetime, n){
  
  x <- matrix(rep(0,n), ncol = n)
  result <- apply(x, 2, busted2)
  return(length(result[result < placetime])/n)
}
sim_busted2(100, 5000)
sim_busted2(500, 5000)
```
```{r}
sim_fivethousand_busted2 <- function(n){
  x <- matrix(rep(0, n), ncol = n)
  result <- apply(x, 2, busted2)
  return(mean(result[result < 5000]))
}
sim_fivethousand_busted2(10000)
```
```{r}
bank_roll2 <- function(x){
  bankroll2 <- 1000
  bet <- 100
  for (i in 1:x){
    ifelse(rbinom(1, 1, 18/38), bankroll <- bankroll - bet, bankroll <- bankroll + bet)
    if (bankroll == 0) break
  }
  return(bankroll)
}

sim_bankroll2 <- function(n, hands){
x <- matrix(rep(hands, n), ncol = n)
result <- apply(x, 2, bank_roll2)
return(c(mean(result), var(result)))
}
sim_bankroll2(5000, 100)
sim_bankroll2(5000, 500)

```

3. **Markov Chains**. Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100.
    a. Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation.
```{r}
markovchain <- function(n){
  win_prob <- 0.48
  bankroll <- 1000
  bet <- 100
  for (i in 1:n){
    if (rbinom(1, 1, win_prob) == 1){
      win_prob <- win_prob + 0.01
      bankroll <- bankroll + bet
    }else{
      win_prob <- 0.48
      bankroll <- bankroll - bet
    }
    if (win_prob >= 1) break

  }
  return(bankroll-1000)
}

sim_markov <- function(n){
  x <- matrix(rep(100000, n), ncol = n)
  result <- apply(x, 2, markovchain)
  return(mean(result))
}

sim_markov(100)
```

    b. Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically?
```{r}
markovchain2 <- function(n){
  win_prob2 <- 0.4892
  bankroll <- 1000
  bet <- 100
  for (i in 1:n){
    if (rbinom(1, 1, win_prob2) == 1){
      win_prob2 <- win_prob2 + 0.01
      bankroll <- bankroll + bet
    }else{
      win_prob2 <- 0.4892
      bankroll <- bankroll - bet
    }
    if (win_prob2 >= 1) break

  }
  return(bankroll-1000)
}

sim_markov2 <- function(n){
  x <- matrix(rep(100000, n), ncol = n)
  result <- apply(x, 2, markovchain2)
  return(mean(result))
}

sim_markov2(100)
```

    c. Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it. 
```{r}
markovchain3 <- function(n){
  win_prob3 <- 0.48
  bankroll <- 1000
  bet <- 100
  for (i in 1:n){
    if (rbinom(1, 1, win_prob3) == 1){
      win_prob3 <- win_prob3 + 0.018
      bankroll <- bankroll + bet
    }else{
      win_prob <- 0.48
      bankroll <- bankroll - bet
    }
    if (win_prob3 >= 1) break
  }
  return(bankroll-1000)
}

sim_markov3 <- function(n){
  x <- matrix(rep(100000, n), ncol = n)
  result <- apply(x, 2, markovchain3)
  return(mean(result))
}

sim_markov3(100)
```


4. For the last two examples in the previous question, you calculated a mean value. Because you saved these final results in a vector, use the bootstrap to estimate the variance of the return in each case for your final answer. Once you have these results, which game has the smaller variance in returns?
```{r}
var.jackknife <- function(a_vector) {
  # Put your function in here
  n <- length(a_vector)
  cita <- rep(NA, n) 
  for (i in 1:n){
    cita[i] <- mean(a_vector[-i])
  }
  jackknife.var <- (n-1)^2/n * var(cita)
  return(jackknife.var)
}


var_sim <- c()
for (i in c("markovchain", "markovchain2", "markovchain3")){
sim_mar <- function(n){
  x <- matrix(rep(100000, n), ncol = n)
  result <- apply(x, 2, i)
  return(result)
}
var_sim <- c(var_sim, var.jackknife(sim_mar(100)))
}
var_sim
```
